---
title: "PFAS Meta-Analysis"
author: "Alexandra Larsen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:  
  html_document:
    toc: yes
    toc_depth: 2
    number_sections: yes
params:
  pfas_type: "PFNA"
  print_code: FALSE
---

<!-- Manuscript: overall, stratified, loo; sensitivity: re-expression, Kwon (PFNA only), 2 vs 3 strata -->
<!-- TODO: link main.Rmd to README.md -->

```{r setup, include = FALSE, root.dir = "C:/Users/alarsen/OneDrive - Environmental Protection Agency (EPA)/Projects/IRIS/PFAS/pfas-birthweight-meta-analysis"}
knitr::opts_chunk$set(echo = params$print_code,
                      warning = FALSE,
                      tidy = TRUE) #tidy.opts = list(width.cutoff = 60), # if word_document
#rmarkdown::render(output_file = paste0("Meta_Analysis_", pfas, ".pdf"))

library(readxl)
library(tidyr)
library(tidyverse)
library(dplyr)
library(metafor)
library(meta)
library(knitr)
library(kableExtra)
library(ggplot2)
library(RColorBrewer)
library(formatR)
source("utility_functions.R")

pfas_type = "PFNA"

if (pfas_type == "PFNA") {
  betas <- read.csv("betas_pfna.csv")
} else if (pfas_type == "PFHxS") {
  betas <- read.csv("betas_pfhxs.csv")
}
attach(betas)
```

# Background

The purpose of this analysis is to investigate the relationship between PFAS exposure during pregnancy and birth weight. Meta-analytic techniques were used to summarize epidemiological findings from `r length(Study)` studies.

# Overall Meta-Analysis

A random effects model was used to estimate the pooled effect of `r pfas_type` exposure on birth weight, capturing any between-study heterogeneity. The model employs weighted estimation with inverse-variance weights.

```{r}
rma_overall <- rma(yi = BetaLnNgml, sei = BetaLnNgmlSE, slab = Study)
print_results("Overall", rma_overall, betas) %>% kable(format = 'pandoc')
```

```{r forest_plot, dev = "png", dpi = 350, fig.width = 7, fit.height = 5}
plot_df <- betas %>%
  #mutate(SampleTiming = fct_recode(Sampling_3, Early = "early", Late = "late", Post = "post")) %>%
  mutate(SampleTiming = fct_recode(Bin, `1st` = "T1", `1st-2nd` = "T1-T2", `2nd` = "T2", 
                                   `2nd-3rd` = "T2-T3", `3rd` = "T3", `Birth` = "B", `Post-Birth` = "PB",
                                   `Pre-Conception` = "PC"),
         Confidence = factor(Confidence, levels = c("High", "Medium", "Low")))
# TODO: reorder factor levels - use levels(); search in old code

forest(rma_overall,
       cex = .55,
       level = 95,
       digits = 1L,
       xlim = c(-1100, 500),
       alim = c(-200, 200),
       order = order(plot_df$Confidence, plot_df$SampleTiming, -plot_df$Sample.Size),
       header = "Study",
       ilab = cbind(as.character(plot_df$Confidence),
                    as.character(plot_df$SampleTiming),
                    plot_df$Sample.Size),
       ilab.xpos = c(-650, -500, -350),
       slab = plot_df$Study,
       xlab = NULL,
       mlab = modify_label(rma_overall, "All Studies, RE Model"))

p <- par(cex = .55, font=2)

text(c(-650, -500, -350), length(Study)+2, c("Confidence", "Timing", "N"))
```

## Leave-One-Out Analyis: Assessing heterogeneity 

```{r}
loo <- leave1out(rma_overall, digits = 2) %>% 
  as.data.frame() %>%
  mutate_at(c(1:3, 5:11), ~round(., 2)) %>%
  mutate_at(c(4), ~round(., 5)) %>%
  add_column(Study, .before = "estimate")

ind <- which(loo$I2 == min(loo$I2))
```

Results indicate that dropping `r loo$Study[ind]` leads to the lowest I2, `r loo$I2[ind]`%. 

<!-- However, the difference between this and the overall $I^2$  is negligible; so we proceed with `r loo$Study[ind]` included. -->
<!-- Sensitivity tests to including Chen are carried out within the medium confidence strata. -->

```{r}
kable(loo, format = 'pandoc')
```

## Publication Bias 

```{r}
# funnel plot code adapted from Kristen Rappazzo's "OZONE and PRETERM BIRTH META-ANALYSIS"
regtest_out <- regtest(rma_overall, model = "rma", predictor = "sei")
regtest_p <- round(regtest_out$pval, 4)
```

The Egger’s regression test did not detect a statistically significant relationship between the observed effect sizes and their standard errors (p = `r regtest_p`), indicating no evidence of funnel plot asymmetry due to publication bias. The funnel plot is displayed below.

```{r funnel_plot, echo = F, include = T, dev = "png"}
funnel(rma_overall, xlab = sprintf("Effect Estimate of ln-%s on Birth-Weight", params$pfas_type)) # label = 3)
```

# Stratified Analyses

In addition to the overall meta-analysis, the studies are also examined according to the following stratification: study confidence, sample timing. 

## Study Confidence

In this section, the studies are examined by confidence levels - low, medium, high, or medium and high combined. Pooled effects are calculated and statistical differences across strata are considered. 

```{r}
conf <- betas[!is.na(Confidence), ] %>% 
  mutate(Confidence2 = factor(ifelse(Confidence == "Low", "Low", "High+Medium"), levels = c("High+Medium", "Low")))
```

```{r}
# Run meta-analysis for each study confidence subgroup using a random effects model
rma_high <- rma(yi = BetaLnNgml, sei = BetaLnNgmlSE, data = conf, subset = Confidence == "High")
rma_med  <- rma(yi = BetaLnNgml, sei = BetaLnNgmlSE, data = conf, subset = Confidence == "Medium")
rma_low  <- rma(yi = BetaLnNgml, sei = BetaLnNgmlSE, data = conf, subset = Confidence == "Low")
rma_hm   <- rma(yi = BetaLnNgml, sei = BetaLnNgmlSE, data = conf, subset = Confidence2 == "High+Medium")
```

```{r}
# Display results
rbind(
  print_results("High", rma_high, conf %>% filter(Confidence == "High")),
  print_results("Medium", rma_med, conf %>% filter(Confidence == "Medium")),
  print_results("Low", rma_low, conf %>% filter(Confidence == "Low")),
  print_results("High+Medium", rma_hm, conf %>% filter(Confidence2 == "High+Medium"))) %>% kable(format = "pandoc")
```

```{r}
# Prepare data for hypothesis test of any significant differences across strata
dat1 <- data.frame(strata = c("High", "Medium", "Low"),
                   estimate = c(coef(rma_high), coef(rma_med), coef(rma_low)),
                   SE = c(rma_high$se, rma_med$se, rma_low$se),
                   tau2 = round(c(rma_high$tau2, rma_med$tau2, rma_low$tau2), 3))

# Run fixed effect model with strata as moderator
res1 <- rma(yi = estimate, sei = SE, mods = ~ strata, method = "FE", data = dat1)

# Save p-value from hypothesis test differences across moderators (rma() runs 
# an omnibus test for moderator variable, i.e., test for any differences across strata)
p1 <- round(res1$QMp, 4)
```

```{r}
# Repeat for High+Medium group
dat2 <- data.frame(strata = c("High+Medium", "Low"),
                   estimate = c(coef(rma_hm), coef(rma_low)),
                   SE = c(rma_hm$se, rma_low$se),
                   tau2 = round(c(rma_hm$tau2, rma_low$tau2), 3))
res2 <- rma(yi = estimate, sei = SE, mods = ~ strata, method = "FE", data = dat2)
p2 <- round(res2$QMp, 4)
```

Hypothesis test results for any significant differences across strata:

* For all three levels, High, Medium and Low: p = `r p1`.
* For High + Medium and Low: p = `r p2`.

## Sample Timing

In this approach, the strata have the following definitions:

* Early: first trimester only, mixture of first and second trimester.
* Late: second trimester only, mixture of second and third trimester, third trimester only.
* Post: at or after birth.

```{r echo = F, warning = F}
rma_early <- rma(yi = BetaLnNgml, sei = BetaLnNgmlSE, mods = ~ Sampling_3, data = betas, subset = Sampling_3 == "early")
rma_late <- rma(yi = BetaLnNgml, sei = BetaLnNgmlSE, mods = ~ Sampling_3, data = betas, subset = Sampling_3 == "late")
rma_post <- rma(yi = BetaLnNgml, sei = BetaLnNgmlSE, mods = ~ Sampling_3, data = betas, subset = Sampling_3 == "post")
# TODO: why does there need to be a mods statement?
```
```{r echo = F, include = T}
rbind(
  print_results("Early", rma_early, betas %>% filter(Sampling_3 == "early")),
  print_results("Late", rma_late, betas %>% filter(Sampling_3 == "late")),
  print_results("Post", rma_post, betas %>% filter(Sampling_3 == "post"))) %>% kable(format = "pandoc")
```
```{r echo = F, echo = F, include = F}
sampling_results <- data.frame(estimate = c(coef(rma_early), coef(rma_late), coef(rma_post)),
                               SE = c(rma_early$se, rma_late$se, rma_post$se),
                               strata = c("T1-T.5", "T2-3", "B-PB"),
                               tau2 = round(c(rma_early$tau2, rma_late$tau2, rma_post$tau2), 3))
rma_sampling <- rma(estimate, sei = SE, mods = ~ strata, method = "FE", data = sampling_results)
p <- round(rma_sampling$QMp, 4)
```

Hypothesis test for any differences across strata: p = `r p`.

# Sensitivity Testing

## Re-expression to Natural Scale

## Two-Strata Approach for the Analysis of Sample Timing

## Continuous Approach for the ANalysis of Sample TIming

 